results

### 2020-12-20 analysis/results.Rmd
sector analysis

```{r}
els_jns_df <- readr::read_csv(here::here("data", "elsevier_apc_list.csv"))
els_jns <- readr::read_csv(here::here("data", "els_historic_jns.csv")) 
els_jns_count <- els_jns %>%
  group_by(year, oa_model) %>%
  count(year) %>%
  group_by(year) %>%
  mutate(prop = n / sum(n)) %>%
  mutate(oa_model_fct = fct_rev(as_factor(oa_model))) 
```

@@

license_df %>%
  group_by(type, license_code) %>%
  summarise(n = sum(articles)) %>%
  mutate(prop = n / sum(n))
all_oa <- license_df %>% 
  group_by(type) %>% 
  summarise(n = sum(articles)) %>%
  pull(n) %>%
  sum()
all_oa
```

@@

```{r}
# invoicing analysis
library(googlesheets4)
tt <- read_sheet("https://docs.google.com/spreadsheets/d/1AqjCLn6VNGJpQ6zdmFSwuhFsDBIITSxijA0r0WLJZTQ/", sheet = 2)
```

```{r}
hybrid_invoice <- hybrid_volume %>%
  filter(oa_sponsor_type == "FundingBody") %>%
 # rename(yrl_volume = articles) %>%
  left_join(tt, by = c("oa_sponsor_name")) %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc-nd"))

sponsor_country <- hybrid_invoice %>%
  # manual changes
  mutate(Country = ifelse(grepl("European|CERN", oa_sponsor_name), "EU", Country)) %>%
  mutate(Country = ifelse(Country == "Multinational", "Other", Country)) %>%
  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
  mutate(country_factor = fct_infreq(country_factor)) %>%
  mutate(country_factor = fct_relevel(country_factor, "Other", after = Inf)) %>%
  group_by(country_factor, `Primary research focus`) %>%
  summarise(n = n_distinct(doi)) %>%
  ungroup() %>%
  mutate(prop = round(n / sum(n), 2))

sponsor_country %>% 
  ungroup() %>%
  pivot_wider(id_cols = c("country_factor"), names_from = `Primary research focus`, values_from = prop) %>%
  janitor::adorn_totals(c("col", "row")) %>%
  rename(`Country/Subject Area` = country_factor) %>%
  knitr::kable()
```

@@

sponsor_country_t <- hybrid_invoice %>%
  # manual changes
  mutate(Country = ifelse(grepl("European|CERN", oa_sponsor_name), "EU", Country)) %>%
  mutate(Country = ifelse(Country == "Multinational", "Other", Country)) %>%
  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
  mutate(country_factor = fct_infreq(country_factor)) %>%
  mutate(country_factor = fct_relevel(country_factor, "Other", after = Inf))

sponsor_country_t %>%
  count(issued_year, Sector, country_factor, cc_norm) %>%
  ggplot(aes(issued_year, n, fill = Sector)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~country_factor)
```

@@

```{r}
hybrid_volume %>% 
  filter(oa_sponsor_type == "ElsevierWaived") %>%
  count(oa_sponsor_name, sort = TRUE) %>%
  mutate(prop = n / sum(n))
```
```{r}
cc_invoice <- hybrid_volume %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc")) %>%
  mutate(oa_sponsor_type = ifelse(is.na(oa_sponsor_type), "unknown", oa_sponsor_type)) %>%
  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
                                  `FundingBody` = "Agreement",
                                  `ElsevierWaived` = "Fee Waived")) %>%
  mutate(fct_source = fct_infreq(oa_sponsor_type)) %>%
  mutate(fct_source = fct_lump(fct_source, prop = 0.05)) %>%
  count(issued_year, fct_source, cc_norm)

>>>>

```{r}
els_jns <- readr::read_csv(here::here("data", "els_historic_jns.csv")) 
els_jns_df <- readr::read_csv(here::here("data", "elsevier_apc_list.csv"))

els_jns_count <- els_jns %>%
  group_by(year, oa_model) %>%
  count(year) %>%
  group_by(year) %>%
  mutate(prop = n / sum(n)) %>%
  mutate(oa_model_fct = fct_rev(as_factor(oa_model))) 
els_jns_count
```

@@

license_df %>%
  group_by(type, license_code) %>%
  summarise(n = sum(articles)) %>%
  mutate(prop = n / sum(n))
all_oa <- license_df %>% 
  group_by(type) %>% 
  summarise(n = sum(articles)) %>%
  pull(n) %>%
  sum()
all_oa
# all articles
license_df %>%
  distinct(all_articles, issued_year) %>%
  pull(all_articles) %>%
  sum()
```

@@

```{r}
# invoicing analysis
library(googlesheets4)
tt <- read_sheet("https://docs.google.com/spreadsheets/d/1AqjCLn6VNGJpQ6zdmFSwuhFsDBIITSxijA0r0WLJZTQ/", sheet = 3)
```

```{r}
hybrid_invoice <- hybrid_volume %>%
  filter(oa_sponsor_type == "FundingBody") %>%
 # rename(yrl_volume = articles) %>%
  left_join(tt, by = c("oa_sponsor_name")) %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc-nd"))
```

```{r license_count}
sponsor_df <- hybrid_invoice %>% 
  filter(!is.na(Country)) %>% 
  # manual changes
  mutate(Country = ifelse(grepl("European", oa_sponsor_name), "EU", Country)) %>%
  mutate(Country = ifelse(Country == "Multinational", "Other", Country)) %>%
  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
  mutate(country_factor = fct_infreq(country_factor)) %>%
  mutate(country_factor = fct_relevel(country_factor, "Other", after = Inf))
sponsor_df %>%
  group_by(country_factor, cc_norm) %>% 
  summarise(n = n_distinct(doi)) %>%
  ggplot(aes(fct_rev(country_factor), n, fill = cc_norm)) +
  geom_bar(stat = "identity", position = position_stack(reverse = TRUE),
            color = "white", alpha = 0.8, size = 0.25) +
  coord_flip() +
  scale_fill_manual(values = 
                      c(`cc-by` = "#B52141",
                        `cc-by-nc-nd` = "#0093c7"),
                    name = "CC license variant") +
  labs(x = "OA Sponsor by Country", y = "Hybrid OA Articles (Agreement)") +
  scale_y_continuous(labels = scales::number_format(big.mark = " "),
                      expand = expansion(mult = c(0, 0.05)),
                     breaks = seq(0,15000, 2000)) +
  theme_minimal_vgrid() +
  theme(legend.position="top", legend.justification = "right")
```

```{r}
sponsor_country <- hybrid_invoice %>%
  filter(!is.na(`Primary research focus`)) %>%
  # manual changes
  mutate(Country = ifelse(grepl("European", oa_sponsor_name), "EU", Country)) %>%
  mutate(Country = ifelse(Country == "Multinational", "Other", Country)) %>%
  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
  mutate(country_factor = fct_infreq(country_factor)) %>%
  mutate(country_factor = fct_relevel(country_factor, "Other", after = Inf)) %>%
  group_by(country_factor, `Primary research focus`) %>%
  summarise(n = n_distinct(doi)) %>%
  ungroup() %>%
  mutate(prop = round(n / sum(n), 2))
```

```{r}
sponsor_country %>%
  ungroup() %>%
  pivot_wider(id_cols = c("country_factor"), names_from = `Primary research focus`, values_from = prop) %>%
  mutate_all(~replace(., is.na(.), 0)) %>%
  janitor::adorn_totals(c("col", "row")) %>%
  rename(`Country/Subject Area` = country_factor) %>%
  knitr::kable()
```

@@

sponsor_country_t <- hybrid_invoice %>%
  # manual changes
  mutate(Country = ifelse(grepl("European", oa_sponsor_name), "EU", Country)) %>%
  mutate(Country = ifelse(Country == "Multinational", "Other", Country)) %>%
  mutate(Country = ifelse(is.na(Country), "Other", Country)) %>%
  mutate(country_factor = fct_lump(Country, prop = 0.025)) %>%
  mutate(country_factor = fct_infreq(country_factor)) %>%
  mutate(country_factor = fct_relevel(country_factor, "Other", after = Inf))

sponsor_df %>%
  count(issued_year, Sector, country_factor) %>%
  mutate(sector_level = factor(Sector)) %>%
  mutate(sector_level = fct_relevel(sector_level, "government", "higher education", "private non-profit", "other")) %>%
  ggplot(aes(gsub("20", "'", issued_year), n, fill = sector_level)) + 
  geom_bar(stat = "identity", position = position_stack(reverse = TRUE),
            color = "white", alpha = 0.8, size = 0.25) +
  facet_wrap(~country_factor) +
  scale_fill_manual("Sector", values = c("#65BADA", "#00A757", "#E5BA52", "grey50")) +
  scale_y_continuous(labels = scales::number_format(big.mark = " "),
                      expand = expansion(mult = c(0, 0.05))) +
  theme_minimal_hgrid() +
  theme(legend.position="top", legend.justification = "right") +
  guides(fill = guide_legend(nrow = 2)) +
  labs(x = "Publication year", y = "Hybrid OA Articles (Agreement)")
```

@@

```{r}
hybrid_volume %>% 
  filter(oa_sponsor_type == "ElsevierWaived") %>%
  count(oa_sponsor_name, sort = TRUE) %>%
  mutate(prop = n / sum(n))
```

```{r}
cc_invoice <- hybrid_volume %>%
  mutate(cc_norm = ifelse(grepl("/by/", URL), "cc-by", "cc-by-nc")) %>%
  mutate(oa_sponsor_type = ifelse(is.na(oa_sponsor_type), "unknown", oa_sponsor_type)) %>%
  mutate(oa_sponsor_type = recode(oa_sponsor_type, 
                                  `FundingBody` = "Agreement",
                                  `ElsevierWaived` = "Fee Waived")) %>%
  mutate(fct_source = fct_infreq(oa_sponsor_type)) %>%
  mutate(fct_source = fct_lump(fct_source, prop = 0.05)) %>%
  count(issued_year, fct_source, cc_norm)
